Doublet splitting of fusion alpha-particle driven ion cyclotron emission

### 2022-02-28 manuscript.tex
tweaked log scale vs linear text

Fig. \ref{fig:2D_F12_physicalpitch} plots solutions for the physically more
realistic pitch-angle
case ($u_\parallel = -0.646 u_0$) in the same manner as Fig.
\ref{fig:2D_F12_zeropitch} except there is an additional black trace in the
upper panel, which represents the measured JET ICE intensity signal; the continuous
trace from Fig. 2 in Ref. \cite{Cottrell1993}.
The observations of ICE intensity are plotted on a log scale scaled to match
the range of the linear MCI growth rates. This juxtaposition is reasonable since
nonlinear PIC code simulations indicate that all growing modes stop when the
fastest growing mode becomes nonlinear at time $\tau$. Under the assumption
that all modes start growing in unison from identical seed amplitudes then the
$log$ of the amplitude of a mode with growth rate $\gamma$ is proportional to
$\gamma \tau$, allowing for a comparison with observations also on a log scale.
There are several points worth noting about this figure. First, the
growth rate is no longer linear with cyclotron harmonic number. Second,
harmonics are broader due to the non-zero parallel drift that gives rise to
strong Doppler shift with increasing $|k_\parallel|$. Third, doublet splitting
of the $10^{th}$ and $11^{th}$ harmonics is observed: see Fig. 2 of Cottrell 1993 for a single
signal showing multiple doublet splittings; Fig. 3 of Dendy et al 1994 (also
Fig. 9 Dendy et al 1995) shows slices at constant propagation angle that when
all superimposed point towards doublet splitting, albeit under-resolved in
wavenumber space. The following describes how these three features are
connected. Doublet splitting arises due to the dip in growth rate along the 
band of instability where
$k_\parallel$ goes through zero combined with the change in sign of the Doppler
shift of the real frequency as $k_\parallel$ changes sign. The effect is
larger for higher harmonics where the growth rate diminishes to zero for
perpendicular propagation.

>>>>

Fig. \ref{fig:2D_F12_physicalpitch} plots solutions for the physically more
realistic pitch-angle
case ($u_\parallel = -0.646 u_0$) in the same manner as Fig.
\ref{fig:2D_F12_zeropitch} except there is an additional black trace in the
upper panel, which represents the measured JET ICE intensity signal; the continuous
trace from Fig. 2 in Ref. \cite{Cottrell1993}.
The observations of ICE intensity are plotted on a log scale scaled to match
the range of the linear MCI growth rates. Direct comparison of log-scale
ICE observations and linear growth rates is justified to the extent that
nonlinear PIC simulations indicate that the linearly unstable modes
cease to grow once the fastest growing mode becomes
nonlinear\cite{Cook2013}. Hence, the $log$ of a mode's amplitude is
proportional to its growth rate under the assumption that
that all modes start growing in unison from identical seed amplitudes.
There are several points worth noting about this figure. First, the
growth rate is no longer linear with cyclotron harmonic number. Second,
harmonics are broader due to the non-zero parallel drift that gives rise to
strong Doppler shift with increasing $|k_\parallel|$. Third, doublet splitting
of the $10^{th}$ and $11^{th}$ harmonics is observed: see Fig. 2 of Cottrell 1993 for a single
signal showing multiple doublet splittings; Fig. 3 of Dendy et al 1994 (also
Fig. 9 Dendy et al 1995) shows slices at constant propagation angle that when
all superimposed point towards doublet splitting, albeit under-resolved in
wavenumber space. The following describes how these three features are
connected. Doublet splitting arises due to the dip in growth rate along the 
band of instability where
$k_\parallel$ goes through zero combined with the change in sign of the Doppler
shift of the real frequency as $k_\parallel$ changes sign. The effect is
larger for higher harmonics where the growth rate diminishes to zero for
perpendicular propagation.
